% Since this a solution template for a generic talk, very little can
% be said about how it should be structured. However, the talk length
% of between 15min and 45min and the theme suggest that you stick to
% the following rules:  

% - Exactly two or three sections (other than the summary).
% - At *most* three subsections per section.
% - Talk about 30s to 2min per frame. So there should be between about
%   15 and 30 frames, all told.


\section{Principles of security}

\subsection{Fundamental design decisions}

\begin{frame}
  \begin{enumerate}
    \item Where to focus security controls?
    \item Where to place security controls?
    \item Complexity or assurance?
    \item Centralised or decentralised control?
    \item Blocking access to the layer below?
  \end{enumerate}
\end{frame}

\begin{frame}
  \begin{block}{Focus and placement of control}
    \begin{itemize}
      \item Focus of control may be on data, operations, or users.

      \item If we look at the control of integrity, its requirements may refer to 
        rules on:
        \begin{itemize}
          \item Format and content of data items, \eg account balance must be 
            integer.

          \item Operations that may be performed on a data item, \eg credit, debit 
            and transfer.

          \item Users who are allowed access to a data item, \eg account holder 
            and bank clerk.
        \end{itemize}
    \end{itemize}
  \end{block}
\end{frame}

%\begin{frame}
%  \begin{block}{Man-Machine Scale}
%    \begin{itemize}
%      \item Applications (focus on users, information)
%      \item Services (middleware)
%      \item Operating system
%      \item Operating system kernel
%      \item Hardware (focus on data, generic)
%    \end{itemize}
%  \end{block}
%\end{frame}

\begin{frame}
  \begin{block}{Complexity or Assurance}
    \begin{itemize}
      \item Generic mechanisms are simple, applications are usually feature rich.

      \item The fundamental dilemma:
        \begin{itemize}
          \item Simple generic mechanisms may not match specific security 
            requirements.

          \item To choose the right features from a rich selection, you need to 
            be a security expert.

          \item Security-unaware users are at a loss.
        \end{itemize}
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
  \begin{block}{Centralized or decentralized controls}
    \begin{itemize}
      \item Within the domain of a security policy, the same controls should be 
        enforced everywhere.

      \item Having a centralized entity to do this makes it easy to achieve 
        uniformity, however, this entity may become a bottleneck.

      \item A distributed solution might be more efficient, however, then you 
        must ensure they all enforce consistently with each other.
    \end{itemize}
  \end{block}
\end{frame}

\subsection{The layer below}

\begin{frame}
  \begin{itemize}
    \item Every security mechanism defines a \emph{security perimeter}.

    \item The parts of a system which can malfunction without breaking the 
      mechanism are said to be outside the perimeter.

    \item The parts of the system that can disable the mechanism are within the 
      perimeter.
  \end{itemize}
\end{frame}

\begin{frame}
  \begin{itemize}
    \item Attackers will try to bypass security mechanisms.

    \item How do you ensure an attacker cannot get access to the layer below 
      the security mechanism?
  \end{itemize}
\end{frame}

\begin{frame}
  \begin{itemize}
    \item Recovery tools, read the sectors directly from the disk; logical 
      access control is implemented in the operating system.

    \item Buffer overruns, a value assigned to a variable is too large for the 
      memory buffer allocated; memory allocated for other variables may be 
      overwritten.

    \item Side-channel analysis, look at the time different operations take to 
      perform, look at power consumption.

    \item JavaScript to perform security checks?
      The client can use a Web page without JavaScript enabled.

  \end{itemize}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[allowframebreaks]
  \small
  \printbibliography
\end{frame}

